\speculake: Hiding Malware in Speculative Execution

### 2018-08-01 paper/related-work.tex
pass on speculative execution subsection of related work

% Lead in 
Leading up to the discovery and release of the Spectre and Meltdown
vulnerabilities multiple teams were working to reverse engineer the mechanisms
that implemented in the microcode of various processors and document their
behavior~\cite{intel-instruction-tables,project_zero,measuring-cache,measuring-rob}.
Research into branch predictors~\cite{godbolt2016branch}, cache replacement
policies, and reorder buffer construction~\cite{measuring-rob} paved the way for
the branch prediction side-channels employed by Spectre (and subsequently
\speculake).
%Efforts to examine cache and memory security also preempted the Spectre work, 
%including Sophia D'Antoine's thesis work developing a side channel based on 
%instruction interleaving in the CPU~\cite{d2015exploiting}. Similar research into branch 
%predictors, cache replacement policies, and reorder buffer construction including work by Matt 
%Godbolt~\cite{godbolt2016branch}, and Henry Wong~\cite{measuring-rob} paved the runway for 
%dedicated branch prodiction side-channels.

% Discovery & Followup
Given the numerous directions that were being investigated in reverse
engineering Intel processors, multiple teams discovered the Spectre
vulnerabilities concurrently~\cite{spectre, project_zero, evtyushkin2018branchscope, maisuradze2018speculose}.
%These works propose multiple variants making use of branch prediction.
Follow 
up work on speculative execution side channels has also demonstrated
web-based vulnerabilites~\cite{genkin2018drive} as well as hardware isolation  
leaks in Intel's SGX~\cite{spectre_sgx}. At the time of writing multiple sources
have claimed new speculative execution related exploits termed \textit{Spectre-NG}
have been reported to Intel and assigned CVE numbers, though details are
not yet available~\cite{spectre_ng_gossip}.
%We have no details
%on content, though Intel has labeled four as critical and another four 
%as intermediate security flaws~\cite{spectre_ng_gossip}.

% defenses

%Much like cache side channel vulnerabilities speculative execution is relatively
%critical to processor performance and requires significant software and hardware 
%updates to fully address. 
% Multiple measures have been proposed to combat the 
% individual variants of Spectre.

% One popular approach uses compiler level software patches to wrap sensitive 
% function calls. Retpoline replaces indirect jumps with \texttt{retq}
% instructions that read from the cached stack (and are thus less likely to be
% mis-speculated)~\cite{retpoline}.
% Gruss~\textit{et al.} propose the KAISER software patch to enforce address space
% isolation for kernel and user space proceses~\cite{gruss2017kaslr}. 
% Other proposals have considered hypervisor enforced time domain separation at varying 
% granularity~\cite{renau2018securing}, and fuzzy clocks that would reduce the accuracy 
% of any time based exfiltration pathway~\cite{hu1992reducing}. However, none of these defenses
% fundamentally defeat the use of a branch predictor as a cooperative mechanism
% spanning multiple processes. Intel has released microarchitecture updates for
% its most recent CPUs, but older architectures do not support these updates and
% will remain vulnerable~\cite{intel-micro-patch}.

% Trippel~\textit{et al.} proposes a method of faithfully simulating hardware
% bugs, given a formal model specifying the
% microarchitecture~\cite{trippel2018meltdownprime}. However, such models are
% unlikely to be released by CPU manufacturers, as they generally involve
% closely-guarded trade secrets, making this a less accessible solution for the
% research community.

%\smallskip
%
%We can certainly look forward to more vulnerabilities relating to speculative 
%branching and related hardware side channels. Our work represents an exploit 
%model for these vulnerabilities reliant upon shared hardware resources.

%%%%%%%%%%%%%%

>>>>

\speculake builds on Spectre~\cite{spectre} and Meltdown~\cite{meltdown} which
leverage speculative execution to leak sensitive information to unprivileged
processes.  Others have built on Spectre and Meltdown to demonstrate web-based
vulnerabilities~\cite{genkin2018drive, schwarz2018netspectre} as well as
hardware isolation leaks in Intel's SGX~\cite{spectre_sgx,chen2018sgxpectre}.
The branch predictor has also been examined~\cite{evtyushkin2018branchscope} in
the light of Spectre but only as side-channel which is leaking information.
\speculake likewise takes advantage of speculative execution however it does so
to attain hidden arbitrary computation. We believe that \speculake can leverage
these advances in speculative execution malware as well.
